---
title: "122_wk11_labActivities2_3"
author: "Margriet Groen"
date: "2024-01-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab activity 2 - Miller and Haden data

# Step 1. Loading the relevant libraries
```{r libraries}
library(broom)
library(tidyverse)
```
# Step 2. Read in the data
```{r read in data}
mh <- read_csv("MillerHadenData.csv")
```

# Step 3.  Construct a scatterplot of the relationship between 'Home' and 'TV'. Also add a line of best fit.

```{r scatterplot Home and TV}
ggplot(mh, aes(x = Home, y = TV)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() +
  labs(x = "Time spend reading at home", y = "Time spend watching TV at home")

```

Question 3a: What can you tell from the scatterplot about the direction of the relationship?

# Step 4.  Conduct the correlation analysis.
```{r correlation}
results <- cor.test(mh$Home, 
                    mh$TV, 
                    method = "pearson", 
                    alternative = "two.sided") %>% 
  tidy()

results
```
```{r pull info}
r <- results %>%
  pull(estimate) %>%
  round(2)
r

df <- results %>%
  pull(parameter)
df

pvalue <- results %>%
  pull(p.value) %>%
  round(3)
pvalue
```

Question 4a:  What is the correlation coefficient (Pearson's r)?

Question 4b: What is the p value?

Question 4c: Is the correlation significant at the p \< .05 level?

Question 4d: What are the degrees of freedom you need to report?

# Step 5. Calculate how much variance in 'time spent reading' can be accounted for by 'time spent watching TV'?

```{r r-squared}
rsquared <- r*r
rsquared
rsquaredPercent <- round(rsquared * 100, 0)
rsquaredPercent
```

# Step 6. Write a few sentences in which you report this result, following APA guidelines.
