---
title: 5. Week 16 -- Hypotheses, associations
subtitle: Written by Rob Davies
order: 5
---

::: callout-warning
This page is under construction: come back soon
:::

```{r}
#| warning: false
#| echo: false
library(kableExtra)
library(tidyverse)
```

```{r}
#| warning: false
#| echo: false
study.one <- read_csv("study-one-general-participants.csv")  
study.two <- read_csv("study-two-general-participants.csv")
```

## Week 16: Introduction {#sec-wk16-introduction}

Welcome to your overview of our work together.

We will complete four classes in **weeks 16-19**.

::: callout-tip
*Putting it all together*

- These classes are designed to help you to revise and to put into practice some of the key ideas and skills you have been developing in the first year research methods modules *PSYC121, PSYC123 and PSYC124*.
- We will do this in the context of a live research project with potential real world impacts: the **Clearly Understood** project.
:::

We will present our PSYC122 lessons in the context of a real research project because we think that this *context* will help you to make sense of the data, and it will help you to see *why* we ask you to practice the skills we are teaching.

We will be revisiting some of the ideas and techniques you have seen introduced in previous classes.
This will give you the opportunity to *revise* and *consolidate* your learning.
We will *extend* your development with some new ideas, to help you to strengthen your skills.

We will observe real scientific findings using the data we will collect together.
These new findings will ultimately show health providers how to communicate their advice more effectively.

### Our learning goals

::: callout-tip
In Week 16, we will ask you to do three things.

- *First* we will ask you to do a pre-lab activity that involves completing a survey.
- *Second* we will ask you to do a set of practical tasks in the lab activity that are designed to consolidate your learning on data **visualization**.
- *Third* we will ask you to **think critically** about potential associations between measures of individual differences and measures of how well people understand health information.
:::

Completing the survey will help you to understand the numbers you will be working with in the activities: where they come from, and what they show about the people tested.
It will also help to teach you about the challenges of measurement: this is one of the key challenges in psychological science.

We will build practice in using correlations to test predictions about associations and *thus* to answer research questions.

## Lectures {#sec-wk16-lectures}

::: callout-tip
Before you go on to the activities, **watch** the lectures:
:::

The lecture for this week is presented in four short parts.
You can view video recordings of the lectures using Panopto, by clicking on the links shown following.

1. **Overview** (15 minutes): What we are doing in weeks 16-20, and *how* and *why* you will develop your critical thinking skills.

```{=html}
<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=219a5ba8-cce8-43e3-afea-afaa010f2d91&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=false&start=9&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player"></iframe>
```

2. **Thinking critically** (11 minutes): A summary of the health communication project, and the ideas we assume to develop our hypotheses.

```{=html}
<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=6fd6d7e9-7457-441d-a689-afaa0113f62d&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=false&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player"></iframe>
```

3. **Thinking about associations** in research (22 minutes): How we visualize and think about distributions and associations.

```{=html}
<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=5bbbe88e-b701-4658-b484-afaa011819e3&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=false&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player"></iframe>
```

4. **Interpreting, reporting and visualizing** correlations (11 minutes): How we use R to estimate and test correlations.

```{=html}
<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=ec9bbeb3-ad95-48e1-a03f-afaa012037f9&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=false&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player"></iframe>
```

::: callout-tip
The slides presented in the videos can be downloaded either as a web page or as a Word document.
:::

- [The slides](data/week16/122-hypotheses-assocations.html) exactly as presented (36 MB). 
- [The slides](data/week16/122-hypotheses-associations-printable-edit.docx) converted to a Word .docx (24 MB). 

You can download the web page `.html` file and click on it to open it in any browser (e.g., Chrome, Edge or Safari). The slide images are high quality so the file is quite big and may take a few seconds to download.

You can download the `.docx` file and click on it to open it as a Word document that you can then edit. 
Converting the slides to a .docx distorts some images but the benefit of the conversion is that it makes it easier for you to add your notes.

## Reading: Links to other classes

We do not provide further reading for this class but you will find it helpful to revise some of the key ideas you have been learning about PSYC122 and in other modules.

- The lectures in *PSYC123* on: the scientific method; reliability and validity; experimental design, especially between-subjects studies; hypothesis testing; and precise hypotheses.
- The lecture in *PSYC122* on correlations. 

## Pre-lab activities {#sec-wk6-prelab-activities}

### Pre-lab activity 1 {#sec-wk6-prelab-activities-1}

In weeks 16-19, we will be working together on a research project to investigate how people vary in their response to health advice.

Completing the project involves collecting responses from *PSYC122* students: **you**.

To enter your responses, we invite you to complete a short survey.

Complete the survey by clicking on the link [here]()

::: callout-tip
In our week 19 class activity, we will analyze the data we collect here.
:::

The survey should take about 20 minutes to complete.

Taking part in the survey is **completely voluntary**.
You can stop at any time without completing the survey if you do not want to finish it.
If you do not want to do the survey, you can do an alternative activity (see below).

All responses will be recorded completely anonymously.

### Pre-lab activity alternative option {#sec-wk6-prelab-activities-2}

If you do not want to complete the survey, we invite you to read the pre-registered research plan for the *PSYC122 health advice* research project.

[Read the project pre-registration](https://osf.io/p6fsc/)

## Lab activities {#sec-wk6-lab-activities}

### Introduction {#sec-wk6-lab-activities-1-introduction}

We will do our practical lab work to develop your skills in the context of the **Clearly Understood** project.

- Our focus will be on what makes it easy or difficult for people to understand written health information.

**Why are we doing this?**

We encounter written health information all the time: in warnings signs, on medication labels, in clinics when we go to see the doctor, and online when we research things we are worried about.
It is not always easy to understand this information.
The problem is that it is unclear how health information *should* be communicated: even this is important, there is not much evidence.

As psychologists, **we can help**.

::: callout-important
In these classes, we will complete a research project to answer the research questions:

1. What person attributes predict success in understanding health information?
2. Can people accurately evaluate whether they correctly understand written health information?
:::

### Get ready {#sec-wk6-prelab-activities-ready}

#### Download the data 

Click on the link: [122_week16_for_students.zip](data/week16/122_week16_for_students.zip) to download the data files folder. Then upload the contents to the new folder you created in RStudio Server.

The downloadable .zip folder includes the data files:

- `study-one-general-participants.csv`
- `study-two-general-participants.csv`

and the R Markdown `.Rmd`:

- `2023-24-PSYC122-w16-how-to.Rmd`

If you can't upload these files to the server, you can use some code to get R to do it for you: uncover the hint to see how.

:::{.callout-tip collapse="true"}
- You can use the code below to directly download the file you need in this lab activity to the server. 
- Remember that you can copy the code to your clipboard by clicking on the 'clipboard' in the top right corner.

1. Get the `study-one-general-participants.csv` data

```{r eval=FALSE}
download.file("https://github.com/lu-psy-r/statistics_for_psychologists/blob/main/PSYC122/data/week16/study-one-general-participants.csv?raw=true", destfile = "study-one-general-participants.csv")
```

2. Get the `study-two-general-participants.csv` data

```{r eval=FALSE}
download.file("https://github.com/lu-psy-r/statistics_for_psychologists/blob/main/PSYC122/data/week16/study-two-general-participants.csv?raw=true", destfile = "study-two-general-participants.csv")
```
:::

#### Check: What is in the data files?

Each of the data files we will work with has a similar structure, as you can see in this extract.

```{r}
#| label: headcheck-wide
#| echo: false
# head(behaviour.rt)
study.two %>%
  filter(participant_ID == c('studytwo.1', 'studytwo.10', 'studytwo.100', 'studytwo.101')) %>%
  as.data.frame() %>%
  kable() %>%
  kable_styling()
```

You can use the *scroll bar* at the bottom of the data window to view different columns.

You can see the columns:

- `participant_ID` participant code
- `mean.acc` average accuracy of response to questions testing understanding of health guidance
- `mean.self` average self-rated accuracy of understanding of health guidance
- `study` variable coding for what study the data were collected in
- `AGE` age in years
- `HLVA` health literacy test score
- `SHIPLEY` vocabulary knowledge test score
- `FACTOR3` reading strategy survey score
- `GENDER` gender code
- `EDUCATION` education level code
- `ETHNICITY` ethnicity (Office National Statistics categories) code

::: callout-tip
It is always a good idea to view the dataset -- click on the name of the dataset in the R-Studio `Environment` window, and check out the columns, scroll through the rows -- to get a sense of what you are working with.
:::

### Lab activity 1: Work with the `How-to` guide {#sec-wk6-lab-activities-1}

The `how-to` guide comprises an .Rmd file:

- `2023-24-PSYC122-w16-how-to.Rmd`

It is full of advice and example code.

The code in the `how-to` guide was written to work with the data file:

- `study-one-general-participants.csv`.

::: callout-tip
We show you how to do everything you need to do in the lab activity (@sec-wk6-lab-activities-2, next) in the `how-to` guide.

- Start by looking at the `how-to` guide to understand what steps you need to follow in the lab activity.
:::

We will take things *step-by-step*.

We split .Rmd scripts by steps, tasks and questions:  

- different steps for different phases of the analysis workflow;
- different tasks for different things you need to do;
- different questions to examine different ideas or coding challenges

::: callout-tip
- Make sure you start at the top of the `.Rmd` file and work your way, in order, through each task.
- Complete each task before you move on to the next task.
:::

In the activity workbook, we are going to work through the following tasks.

**Step 1: Set-up**

1. Empty the R environment -- using `rm(list=ls())`
2. Load relevant libraries -- using `library()`

**Step 2: Load the data**

3. Read in the data file -- using `read_csv()`
4. Inspect the data -- using `head()` and `summary()`
5. Change the type classification of a variable in the data -- using `as.factor()`

**Step 3: Use histograms to examine the distributions of variables**

6. Draw histograms to examine the distributions of variables -- using `ggplot()` and `geom_histogram()`

**Step 4: Edit your plots to make them look good**

7. Edit the appearance of one histogram plot step-by-step

**Step 5: Now draw scatterplots to examine associations between variables**

8. Draw scatterplots to examine the associations between some variables -- using `ggplot()` and `geom_point()`
9. Draw scatterplots to examine different variables

**Step 6: Edit the scatterplots to make them look good**

10. Edit the appearance of a plot step-by-step

**Step 7: Use correlation to to answer the research questions**

11. Examine associations between variables using correlation.

::: callout-tip
If you are unsure about what you need to do, look at the advice in `2023-24-PSYC122-w16-how-to.Rmd` on how to do the tasks, with examples on how to write the code.
:::

You will see that you can match a task in the `activity` workbook to the same task in the `how-to` guide.
The `how-to` shows you what function you need and how you should write the function code.

This process of adapting demonstration code is a process critical to data literacy and to effective problem solving in modern psychological science.

::: callout-warning
Don't forget: You will need to change the names of the dataset or the variables to complete the tasks in the activity.
:::

### Lab activity 2 {#sec-wk6-lab-activities-2}

#### Questions {#sec-wk6-lab-activities-1-questions}

## Answers {#sec-wk6-lab-activities-answers}

When you have completed all of the lab content, you may want to check your answers with our completed version of the script for this week. **Remember**, looking at this script (studying/revising it) does not replace the process of working through the lab activities, trying them out for yourself, getting stuck, asking questions, finding solutions, adding your own comments, etc. **Actively engaging** with the material is the way to learn these analysis skills, not by looking at someone else's completed code...

## Online Q&A {#sec-wk6-lab-Q-and-A}

Below is the recording of this week's online Q&A.

```{=html}
```
